\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-09 paper/algorithm.tex
paper: added query schematic for basic plans
paper: keep refining algorithm section

\subsection{Query planning}
The input to our query planner is a tuple $(T, \Phi, \Psi, P, G, A)$: a set of tables $T$, a set of filter predicates $\phi_t, t \in T$, a set of join predicates $\psi_{(t_1, t_2)}, t_1, t_2 \in T$, a set of attributes $P$, and an optional set of attributes $G$ and aggregation function $A \in \{\text{Max}, \text{Min}, \text{Sum}, \text{Avg}, \text{Count}\}$.

@@

To reduce the search space, we only consider the minimal imputation, the maximal imputation, and the minimal drop. The minimal imputation (resp. drop) only imputes (resp. drops) the columns required by the relational operator immediately following the imputation. The maximal imputation imputes all columns in the relation, regardless of which are required.  Algorithm~\ref{algo:plan-helpers} presents
a series of helper functions used by the top level planner, shown in Algorithm~\ref{algo:top-level-planner}.

>>>>

\subsection{Query planning}
The input to our query planner is a tuple $(T, F, J, P, G, A)$:
\begin{itemize}
\item A set of tables $T$.
\item A relation $F: T \times \Phi$ between tables and filter predicates.
\item A relation $J : T \times \Psi \times T$ between tables and join predicates.
\item A set of projection attributes $P$.
\item An optional set of grouping attributes $G$ and aggregator function $A$.
\end{itemize}

@@

Before describing the operation of the query planner, we describe the semantics of a specialized data structure for holding sub-plans, called a \emph{plan cache}.
At a high level, a plan cache is a mapping from a set of tables to a set of optimal plans that perform the required joins and filters over these tables.
We store multiple plans rather than a single best plan because each plan has a distinct set of dirty attributes.

\begin{figure}
  \begin{align*}
    \llbracket Q[T] \rrbracket &= \begin{cases}
      P & (T, P) \in Q \\
      \emptyset & \text{otherwise} \\
    \end{cases} \\
    \llbracket Q[T] \lhd p \rrbracket &= \begin{aligned}
                                           Q \cup \{(T,\ \argmin_{q \in P}\ \textsc{Cost}(q))\}\ \text{where} \\
                                           P = \left\{q ~|~ \begin{aligned}
                                                                     q \in \llbracket Q[T] \rrbracket \cup \{p\}, \\
                                                                     \textsc{Dirty}(q) = \textsc{Dirty}(p) 
                                                                   \end{aligned}
                                           \right\}
                                         \end{aligned}
    % \in 
  \end{align*}
  \caption{Semantics of the plan cache.}
\end{figure}

The query planner (Algorithm~\ref{algo:top-level-planner}), operates as follows.
First, it collects the set of attributes which are used by the operators in the plan or which are visible in the output of the plan.
This set will be used to determine which attributes are imputed.
Then, it constructs a plan cache

To reduce the search space, we only consider the minimal imputation, the maximal imputation, and the minimal drop. The minimal imputation (resp. drop) only imputes (resp. drops) the columns required by the relational operator immediately following the imputation. The maximal imputation imputes all of the attributes in the relation which are used by other operators in the query plan or are visible in the query output.

Algorithm~\ref{algo:plan-helpers} presents a series of helper functions used by the top level planner, shown in Algorithm~\ref{algo:top-level-planner}.

### 2017-02-09 paper/algorithms/plan-helpers.tex
paper: keep refining algorithm section

%\begin{algorithm}
  \begin{algorithmic}
    \Require{$q$ is a query plan. $C_{req}$ is the set of attributes that must be imputed in the output of this query plan. $C_{vis}$ is the set of attributes which are visible in the output of the overall query.}
    \Ensure{A set of query plans $Q$ with added imputation.}
    \Function{AddImpute}{$q, C_{req}, C_{vis}$}
    \State $D_{must} \gets \Call{Dirty}{q} \cap C_{req}$
    \State $D_{may} \gets D_{must} \cup (\Call{Dirty}{q} \cap C_{vis})$
    \State $Q \gets \emptyset$
    \If{$D_{must} = \emptyset$}
    \State $Q \gets Q \cup \{q\}$
    \Else
    \State $Q \gets Q \cup \{\mu_{D_{must}}(q),\ \delta_{D_{must}}(q)\}$
    \EndIf

@@

    \Require{$t$ is a table and $\phi$ is a filter predicate.}
    \Ensure{A set of plans for filtering $t$ with $\phi$.}
    \Function{OptFilter}{$t, \phi, C_{vis}$}
    \State \Return $\{\sigma_{\phi}(q)\ |\ q \in \Call{AddImpute}{t, \Call{Attrs}{\phi}, C_{vis}}\}$
    \EndFunction

@@

    \Require{$q$ is a query plan, $G$ is a set of grouping attributes, $A$ is an aggregation function, and $C_{vis}$ is a set of visible attributes.}
    \Ensure{A set of plans for grouping \& aggregating with $G$ and $A$.}
    \Function{OptGroupBy}{$q, G, A, C_{vis}$}
    \State $C \gets G \cup \Call{Attrs}{A}$
    \State $Q' \gets \Call{AddImpute}{q, C, C_{vis}}$
    \State \Return $\{\Call{GroupBy}{q', G, A} ~|~ q' \in Q'\}$
    \EndFunction

@@

    \Require{A set of tables $T$, a plan cache $Q$ containing at least one query plan for each table in $T$, and a relation $J : T \times \Psi \times T$ between tables and join predicates.}
    \Ensure{An updated plan cache $Q'$ which contains at least one plan that combines all of the tables in $T$.}
    \Function{OptJoin}{$Q, T, J, C_{vis}$}
    \For{$size \in 2...|T|$}
    \For{$S \in \{\text{all length}\ size\ \text{subsets of}\ T\}$}
    \For{$t \in S$}
    \State $S' \gets S \setminus \{t\}$
    \For{$(t, \psi, t') \in J$ where $t' \in S'$}
    \State $C \gets \Call{Attrs}{\psi}$
    \State $L \gets \{\Call{AddImpute}{q, C, C_{vis}} ~|~ q \in Q[t] \}$
    \State $R \gets \{\Call{AddImpute}{q, C, C_{vis}} ~|~ q \in Q[S'] \}$
    \State $Q[S] \lhd \{l \bowtie_\psi r ~|~ l \in L, r \in R\}$
    \EndFor
    \EndFor
    \EndFor
    \EndFor
    \EndFunction
  \end{algorithmic}
  \caption{Base algorithms for query planning with imputations.}
  \label{algo:plan-helpers}
%\end{algorithm}

>>>>

%\begin{algorithm}
  \begin{algorithmic}
    \Require{$q$ is a query plan. $C_l$ is a set of attributes that must be imputed in the output of this query plan. $C_g$ is the set of attributes which are used in the final plan.}
    \Ensure{A set of query plans $Q$ with added imputation.}
    \Function{AddImpute}{$q, C_l, C_g$}
    \State $D_{must} \gets \Call{Dirty}{q} \cap C_l$
    \State $D_{may} \gets D_{must} \cup (\Call{Dirty}{q} \cap C_g)$
    \State $Q \gets \emptyset$
    \If{$D_{must} = \emptyset$}
    \State $Q \gets Q \cup \{q\}$
    \Else
    \State $Q \gets Q \cup \{\mu_{D_{must}}(q),\ \delta_{D_{must}}(q)\}$
    \EndIf

@@

    \Require{$t$ is a table and $\phi$ is a filter predicate.}
    \Ensure{A set of plans for filtering $t$ with $\phi$.}
    \Function{OptFilter}{$t, \phi, C_g$}
    \State \Return $\{\sigma_{\phi}(q)\ |\ q \in \Call{AddImpute}{t, \Call{Attrs}{\phi}, C_g}\}$
    \EndFunction

@@

    \Require{$q$ is a query plan, $G$ is a set of grouping attributes, $A$ is an aggregation function, and $C_g$ is a set of visible attributes.}
    \Ensure{A set of plans for grouping \& aggregating with $G$ and $A$.}
    \Function{OptGroupBy}{$q, G, A, C_g$}
    \State $C \gets G \cup \Call{Attrs}{A}$
    \State $Q' \gets \Call{AddImpute}{q, C, C_g}$
    \State \Return $\{\Call{GroupBy}{q', G, A} ~|~ q' \in Q'\}$
    \EndFunction

@@

    \Require{A set of tables $T$, a plan cache $Q$ containing at least one query plan for each table in $T$, and a relation $J : T \times \Psi \times T$ between tables and join predicates.}
    \Ensure{An updated plan cache $Q'$ which contains at least one plan that combines all of the tables in $T$.}
    \Function{OptJoin}{$Q, T, J, C_g$}
    \For{$size \in 2...|T|$}
    \For{$S \in \{\text{all length}\ size\ \text{subsets of}\ T\}$}
    \For{$t \in S$}
    \State $S' \gets S \setminus \{t\}$
    \For{$(t, \psi, t') \in J$ where $t' \in S'$}
    \State $C \gets \Call{Attrs}{\psi}$
    \State $L \gets \{\Call{AddImpute}{q, C, C_g} ~|~ q \in Q[t] \}$
    \State $R \gets \{\Call{AddImpute}{q, C, C_g} ~|~ q \in Q[S'] \}$
    \State $Q[S] \lhd \{l \bowtie_\psi r ~|~ l \in L, r \in R\}$
    \EndFor
    \EndFor
    \EndFor
    \EndFor
    \EndFunction
  \end{algorithmic}
  \caption{Base algorithms for query planning with imputations.}
  \label{algo:plan-helpers}
%\end{algorithm}
